---
title: "R-Exercise1"
author: "Riyanshi Bohra"
format: html
editor: visual
---

## Homework 1: In R

Installing the required packages:

```{r}
if (!require("pacman"))
 install.packages("pacman")
```

```{r}
#Loading the libraries
library(pacman)

p_load(dlookr,
       DMwR2, # Data Mining with R functions
       GGally, # Pair-wise plots using ggplot2
       Hmisc, # Data analysis 
       palmerpenguins, # Alternative to the Iris dataset
       tidyverse) # Data wrangling, manipulation, visualization
```

Loading the data:

```{r}
data(algae, package = "DMwR2")
algae |> glimpse()
```

Finding the Mean, Median and Mode: Central Tendency-

```{r}
# Calculating Mean
algae$a1 |>
  mean()
```

```{r}
# Calculating Median
algae$a1 |>
  median()
```

```{r}
#Calculating Mode
Mode <- function(x, na.rm=FALSE){
if(na.rm) x<-x[!is.na(x)]
ux <- unique (x)
return (ux[which.max(tabulate(match(x, ux)))])
}

algae$a2 |> Mode()
```

Using the DMwR centralValue() function:

```{r}
# Numerical variable
algae$a1 |> centralValue()
# Nominal variable
algae$speed |> centralValue()
```

Finding the Statistics of spread: Variation-

```{r}
# Finding Variance
algae$a1 |> var()
```

```{r}
# Finding the Standard Deviation
algae$a1 |> sd()
```

```{r}
# Finding the Range
algae$a1 |> range()
```

```{r}
# Finding the maximum value
algae$a1 |> max()
```

```{r}
# Finding the minimum value
algae$a1 |> min()
```

```{r}
# Finding the interquartile range
algae$a1 |> IQR()
```

```{r}
# Finding the Quantiles
algae$a1 |> quantile()
algae$a1 |> quantile(probs = c(0.2, 0.8)) #Specifying specific quantiles
```

Finding the Missing Values:

```{r}
# Calculating the total number of null values in the dataset

library(purrr)
nas <- algae %>% 
  purrr::map_dbl(~sum(is.na(.))) %>% 
  sum()

cat("The dataset contains ", nas, "NA values. \n")
```

```{r}
# Calculating the total number of incomplete rows in the dataset

incomplete_rows <- algae %>% 
  summarise_all(~!complete.cases(.)) %>%
  nrow()

cat("The dataset contains ", incomplete_rows, "(out of ", nrow(algae),") incomplete rows. \n") #printing the number of incomplete rows
```

Summarizing the dataset:

```{r}
# Using the summary() 
algae |> summary()
```

```{r}
# Using the Hmisc's describe() 
data("penguins")
penguins |> Hmisc::describe()
```

```{r}
# Using the dlookr's describe() 
penguins |> dlookr::describe()
```

Finding the summary of a subset of data:

```{r}
algae |>
  summarise(avgNO3 = mean(NO3, na.rm=TRUE),
            medA1 = median(a1))
```

```{r}
# Using the summarise_all() method
algae |>
  select(mxPH:Cl) |>
  summarise_all(list(mean, median), na.rm = TRUE)
```

```{r}
algae |>
  select(a1:a7) |>
  summarise_all(funs(var))
```

```{r}
algae |>
  select(a1:a7) |>
  summarise_all(c("min", "max"))
```

Using the summarise() with group_by():

```{r}
algae |>
  group_by(season, size) |>
  summarise(nObs = n(), mA7 = median(a7))
```

```{r}
penguins |> 
  group_by(species) |>
  summarise(var = var(bill_length_mm, na.rm = TRUE))
```

Aggregating the data:

```{r}
# Using reframe()
penguins |>
  group_by(species) |>
  reframe(var = quantile(bill_length_mm, na.rm = TRUE))
```

```{r}
# Using dlookr
penguins |>
  group_by(species) |>
  dlookr::describe(bill_length_mm)
```

Advanced:

1.  List data types of the attributes in your tidy dataset:

```{r}
str(algae)
```

2.  Check for skewness in data distribution in the attributes:

```{r}
algae %>% 
  select_if(is.numeric) %>%
  summarise(across(everything(), skewness, na.rm = TRUE))
```

3.  Check for correlations among the attributes:

```{r}
# Finding the correlations
cor_matrix <- cor(algae %>% select_if(is.numeric), use = "complete.obs")
```

```{r}
# Installing and importing necessary libraries for visualizing the correlations between attributes 
if(!require("corrplot"))
  install.packages("corrplot")
```

```{r}
library(corrplot)
```

```{r}
# Visualizing correlations using corrplot
corrplot(cor_matrix, method = "circle", type = "lower", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

4.  Examine the extent of missing data and the best way to deal with them:

```{r}
#Examining the extent of missing data:
missing_data_summary <- algae %>% 
  summarise(across(everything(), ~mean(is.na(.))*100))
print(missing_data_summary)
```

```{r}
# Suggesting ways to deal with missing data
cat("Best ways to deal with missing data are: \n")
cat("1. Using mean, median, or mode for the replacement of missing values in numeric variables.\n")
cat("2. If the amount of missing data is relatively low, we can consider eliminating those rows or columns. \n")
cat("3. Using algorithms like KNN or regression to impute missing values.\n")
```
